Biodiversity improves tropical forest resilience

### 2019-01-08 Main Document.Rmd
Draft 0.1.9013 Main document & Cover letter

---
title: Biodiversity improves tropical forest resilience
author: Sylvain Schmitt$^{1,*}$, Isabelle Maréchaux$^2$, Jerome Chave $^3$, Fabian Fischer$^3$, Camille  Piponiot$^4$, Stéphane Traissac$^5$, Bruno Hérault$^6$
output:
  bookdown::word_document2: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: no
    toc: no
    keep_tex: yes
  bookdown::html_document2:
    number_sections: no
    toc: yes
    toc_float: yes
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
theme_set(bayesplot::theme_default())
opts_chunk$set( echo = F, message = F, warning = F, 
                fig.height = 8, fig.width = 12,
                cache = T, cache.lazy = F)
```

@@

$^1$ UMR BIOGECO, Université de Bordeaux, 69 route d’Arcachon, F-33612 Cestas cedex France

$^2$ AMAP, INRA, University of Montpellier, IRD, CIRAD, CNRS, , Montpellier, France

$^3$ Laboratoire Evolution et Diversité Biologique, UMR5174, CNRS, Université Paul Sabatier, IRD, Toulouse Cedex 9, France

$^4$ Université de Guyane, UMR EcoFoG (Agroparistech, CNRS, Inra, Université des
Antilles, Cirad), Kourou, French Guiana

$^5$ AgroParisTech, UMR ECOlogie des FOrêts de Guyane, 97387 Kourou, France

$^6$ INPHB, Institut National Polytechnique Félix Houphouët-Boigny, BP 1093, Yamoussoukro, Ivory Coast

$^*$ Corresponding author: sylvain.schmitt@agroparistech.fr

@@

@@

```{r, eval=F}
doe %>% 
  select(starts_with("CWM")) %>% 
  reshape2::melt() %>% 
  group_by(variable) %>% 
  summarise_all(funs(mean, sd, min, max))
```

@@

The recovery of the simulated forest at a time $t$, $R(t)$ regarding a given property $X$ was defined as follows (Figure \@ref(fig:IrGraphFinal)):

@@

When $R(t)$ approaches $R_{eq}=1$, the disturbed and control systems are considered equivalent, indicating a perfect recovery. We then calculated the Euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$ (Figure \@ref(fig:IrGraphFinal)). Because we set up the intensity of disturbance, our analysis did not include any ecosystem resistance, i.e. the ecosystem capacity to absorb a disturbance event. Consequently, in this particular case, we used the inverse of the cumulative Euclidean distance to the equilibrium as an index of ecosystem resilience, further defined as $I_R$ (Figure \@ref(fig:IrGraphFinal)):

@@

<!-- $$R[X(t)] = \frac{X_D(t)}{X_C(t)}$$ -->
<!-- $$d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$$ -->
<!-- $$I_R(X) = \frac1{\int _{t=disturb_{iter}} ^{t=disturb_{iter}+600} d_{eq}(t).dt}$$ -->

@@

```{r, eval=F}
doe %>% 
  select(SR, FDiv, FEve) %>% 
  reshape2::melt(id.vars = "SR") %>% 
  mutate(variable = paste0("$", variable, "$")) %>% 
  group_by(variable, SR) %>% 
  summarise(mean =  paste0("$", round(mean(value),2), 
                           "\\pm", round(sd(value),2), "$")) %>% 
  arrange(variable, desc(SR)) %>% 
  kable(caption = 'Mean and standard deviation of the predisturbance functional diversity indices for each level of species richness.',
        format = 'pandoc', col.names = c("Variable", "Species richness ($SR$)", 
                                         "Mean")) %>% 
  kable_styling(full_width = F)
```

As species richness ($SR$) increased from 5 to 125 species, functional diversity varied from $0.76\pm0.18$ to $0.96\pm0.02$ and evenness ($FEve$) varied from $0.46\pm0.24$ to $0.64\pm0.02$. Species richness ($SR$) was significantly correlated to functional indices (one-way ANOVA $p-value < 0.001$ for both $FDiv$ and $FEve$) and model $AIC$ was higher without $SR$, consequently we kept only $FDiv$ and $FEve$ within linear regressions. Both functional diversity ($FDiv$) and functional evenness ($FEve$) were positively related with $I_R$, i.e. forest resilience increased with both functional diversity and functional evenness (Figure \@ref(fig:gIr) and Table \@ref(tab:tgIr)). Similar results were obtained for all disturbance levels, with smaller p-values for higher disturbance intensity.

```{r gIr, echo=FALSE, fig.cap='Variation of forest resilience with taxonomic and functional diversity for different levels of disturbance. Resilience index $I_R$ was represented against functional diversity ($FDiv$) for different levels of disturbance ($disturb_{intensity}$} = 25, 50 and 75% of total basal area); dot shapes varies with species richness whereas dot color indicates functional evenness ($FEve$). Solid lines and grey areas show fitted relationships with their mean and variance of the form $I_R = FDiv + \\epsilon$.', fig.width=10, fig.height=15}
Ir %>% 
  left_join(doe, by = "sim") %>% 
  ungroup() %>% 
  select(disturbance, 
         structure, SR, FEve, FDiv) %>% 
  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), value.name = "Ir") %>% 
  ggplot(aes(FDiv, Ir, shape = SR, col = FEve, group = NA)) +
  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
  geom_point(size = 4) +
  facet_wrap(~disturbance, scales = "free", labeller = 'label_both', 
             nrow = 3) + 
  xlab('Functional diversity (FDiv)') +
  ylab('Resilience indice Ir') +
  scale_shape_discrete('Species\nrichness\n(SR)') +
  scale_color_continuous('Functional\nevenness\n(FEve)') +
  theme(text = element_text(size=20),
        legend.key.size=unit(4,"line"))
```

```{r tgIr, cache=T}
Ir %>% 
  left_join(doe, by = "sim") %>% 
  ungroup() %>% 
  mutate(SR = as.numeric(SR)) %>% 
  select(disturbance, structure, production,
         SR, FEve, FDiv) %>% 
  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
                 variable.name = "Ir") %>% 
  group_by(Ir, disturbance) %>%
  do(fit = lm(log(value) ~ FEve + FDiv, data = .)) %>% 
  # do(fit = lm(log(value) ~ FEve + FDiv, data = .)) %>% 
  tidy(fit) %>% 
  ungroup() %>% 
  filter(term != "(Intercept)") %>% 
  filter(!grepl("SR", term)) %>% 
  mutate(`P-value` = ifelse(p.value < 0.001, "***",
                            ifelse(p.value < 0.01, "**",
                                   ifelse(p.value < 0.05, "*",
                                          ifelse(p.value < 0.1, ".", 
                                                 "n.s."))))) %>% 
  mutate(Coefficient = round(estimate,2)) %>% 
  mutate(value = paste(Coefficient, `P-value`)) %>%
  mutate(value = cell_spec(value,
                           bold = ifelse(grepl("*", value,
                                               fixed = T), T, F))) %>%
  reshape2::dcast(Ir + disturbance ~ term, value.var = "value") %>% 
  left_join(Ir %>% 
              left_join(doe, by = "sim") %>% 
              ungroup() %>% 
              mutate(SR = as.factor(SR)) %>% 
              select(disturbance, structure, production,
                     SR, FEve, FDiv) %>% 
              reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
                             variable.name = "Ir") %>% 
              group_by(Ir, disturbance) %>%
              do(fit = lm(log(value) ~ FEve + FDiv, data = .)) %>% 
              glance(fit) %>% 
              mutate(r.squared = round(r.squared, 2)) %>% 
              select(Ir, disturbance, r.squared)) %>% 
  mutate(Ir = ifelse(Ir == "production", "functioning", "structure")) %>% 
  kable(caption = "Effects of functional diversity on forest resilience. Results of the linear regression testing the effect of functional diversity indices (FDiv and FEve) on the resilience index ($I_R$) for forest structure and functioning and for each level of disturbance intensity ($disturb_{intensity}$). To meet the assumption of normality $I_R$ was log-transformed. Test of FDiv and FEve effects was done by linear regression with FDiv and FEve as independent variables. Values of FDiv and FEve are coefficients from linear regressions. Significant $p-values$ are shown in bold. ns, non-significant; . p < 0.1; * p < 0.05; ** p < 0.01; *** p < 0.001.",
        escape = F, col.names = c("$I_R$", "Disturbance intensity",
                                  "FDiv", "FEve", "$R^2$")) %>% 
  kable_styling(full_width = F) %>% 
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1:2)
```

```{r anova, eval=F}
tHSD <- doe %>% 
  ungroup() %>% 
  mutate(SR = as.factor(SR)) %>% 
  select(disturbance,
         SR, FEve, FDiv) %>% 
  reshape2::melt(id.vars = c("disturbance", "SR")) %>% 
  group_by(variable) %>% 
  do(anova = aov(value ~ SR, data = .)) %>% 
  mutate(tHSD = TukeyHSD(anova, ordered = F, conf.level = 0.95)) %$%
  lapply(tHSD, function(x) multcompView::multcompLetters(x[,4])) %>% 
  lapply(function(x) data.frame(SR = names(x$Letters), group = x$Letters)) %>% 
  bind_rows() %>% 
  mutate(axis = c(rep("FEve", 3), rep("FDiv", 3)))
doe %>% 
  ungroup() %>% 
  mutate(SR = as.factor(SR)) %>% 
  select(disturbance,
         SR, FEve, FDiv) %>% 
  reshape2::melt(id.vars = c("disturbance", "SR")) %>% 
  left_join(tHSD) %>% 
  ggplot(aes(SR, value, fill = SR, label = group)) +
  geom_boxplot() +
  geom_text(aes(y = 1.1)) +
  ggpubr::stat_compare_means(method = "anova", aes(label = paste("Anova", ..p.signif..))) +
  facet_wrap(~ variable, scales = "free") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

```{r residualsHist, eval=F}
Ir %>% 
  left_join(doe, by = "sim") %>% 
  ungroup() %>% 
  mutate(SR = as.factor(SR)) %>% 
  select(disturbance, structure, production,
         SR, FEve, FDiv) %>% 
  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
                 variable.name = "Ir") %>% 
  group_by(Ir, disturbance) %>%
  do(residuals = residuals(lm(log(value) ~ FEve + FDiv, data = .))) %>% 
  tidy(residuals) %>% 
  ggplot(aes(x)) +
  geom_histogram(bin = 20) +
  geom_vline(xintercept = 0, col = "red") +
  facet_grid(~ Ir + disturbance, scales = "free") +
  xlab("Residuals")
```

```{r residualsTest, eval=F}
Ir %>% 
  left_join(doe, by = "sim") %>% 
  ungroup() %>% 
  mutate(SR = as.factor(SR)) %>% 
  select(disturbance, structure, production,
         SR, FEve, FDiv) %>% 
  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
                 variable.name = "Ir") %>% 
  group_by(Ir, disturbance) %>%
  do(residuals = residuals(lm(log(value) ~ FEve + FDiv, data = .))) %>% 
  tidy(residuals) %>% 
  group_by(Ir, disturbance) %>% 
  do(normality = shapiro.test(.$x)) %>%
  tidy(normality) %>% 
  transmute(p.value = ifelse(p.value < 0.001, "***",
                             ifelse(p.value < 0.01, "**",
                                    ifelse(p.value < 0.05, "*",
                                           ifelse(p.value < 0.1, ".", 
                                                  "n.s.")))))
```

@@

The net effect of biodiversity on each ecosystem property was quantified and partitioned into selection and complementarity effects after normalizing, as explained in the Methods section. We found that the complementarity effect sharply increased immediately after disturbance, was the dominant effect of biodiversity during the first decades and then progressively vanished (Figure \@ref(fig:gBE)). When increasing the disturbance intensity, the effect of complementarity increased further and dominated even longer. The selection effect, on the other hand, sharply decreased just after disturbance, and only gradually came to play a more important role, dominating the net biodiversity effect ca. a century after disturbance. Finally, at high disturbance intensity, the net effect of biodiversity on aboveground biomass was still lower than the control 600 years after disturbance (Figure \@ref(fig:gBE)). We found similar results, but with an amplified signal, for basal area ($BA$) and stem abundance ($N$). Finally, forest growth primary productivity ($GPP$) recovered only within a few years with a dominant complementarity effect.

```{r gBE, echo=FALSE, fig.cap='Post-disturbance trajectories of complementarity and selection effects on aboveground biomass for increasing levels of disturbance (25%, 50% and 75% of basal area loss). The net biodiversity ($NE$), complementarity ($CE$) and selection ($SE$) effects were normalized by the net effect of biodiversity in the control simulation ($disturb_{intensity}=0\\%$). Solid lines represents the mean value whereas shaded areas represent the range of values across the 20 simulations with a species richness ($SR$) of 125. Time axis has been square-root transformed to highlight first decades dynamic.'}
BEresilience %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  # filter(BE %in% c('SE', 'CE')) %>%
  group_by(time, BE, disturbance, SR) %>%
  summarise(qt1 =quantile(agb, probs=0.05),
            qt2 =quantile(agb, probs=0.75),
            mean=mean(agb),
            n=n()) %>% 
  ggplot(aes(x = time)) +
  geom_hline(yintercept = 1, color = 'red') +
  geom_hline(yintercept = 0, color = 'blue') +
  geom_ribbon(aes(ymin = qt1, ymax = qt2, fill = BE), alpha = 0.3) +
  geom_line(aes(y = mean, color = BE)) +
  facet_grid(disturbance ~ SR, labeller = 'label_both') +
  xlab('Time since disturbance (years)') +
  ylab('Biodiversity effect on aboveground biomass') +
  scale_x_sqrt() +
  theme(text = element_text(size=20),
        legend.key.size=unit(2,"line"))
```

@@

```{r LiteratureBiodiversityEffect, eval=F}
literature <- data.frame(
  Study = c("@Bauters2015", "@Finegan2015", "@Prado-Junior2016",
            "@Lohbeck2016", "@DeAvila2018", "@Marechaux2017", "@Tobner2016",
            "@Morin2014a", "@Cardinale2007", "@VanderSande2017a",
            "@Poorter2017", "@Huang2018", "@Niklaus2017a",
            "@Fargione2007"),
  StudyType = c("Observation", "Observation", "Observation",
                "Observation", "Experiment", "Simulation", "Experiment",
                "Simulation", "Review", "Review",
                "Observation", "Experiment", "Experiment", 
                "Experiment"),
  Type = c("Forest", "Forest", "Forest", 
           "Forest", "Forest", "Forest", "Forest",
           "Forest", "Grasslands-Tundra-Estuaries", "Forest",
           "Forest", "Forest", "Forest", 
           "Grassland"),
  Biome = c("Tropical", "Tropical", "Tropical",
            "Tropical", "Tropical", "Tropical", "Temperate",
            "Temperate", "Temperate", "Tropical",
            "Tropical", "SubTropical", "SubToprical", 
            ""),
  SubBiome = c("Wet", "Wet-Dry", "Dry",
               "Wet?", "Wet", "Wet", "",
               "", "", "",
               "Wet", "", "", 
               ""),
  Effect = c("SE", "SE", "SE",
             "SE", "SE", "SE", "SE",
             "CE", "CE", "CE",
             "CE", "CE", "CE", 
             "SE-CE"),
  Disturbed = c("Y", "N?", "Y",
                "Y", "Y", "N", "Y",
                "N", "?", "?",
                "50%Y", "Y", "", 
                "Y"),
  Age = c("77yo", "", "12yo",
          "1-29yo", "21yo", "600yo", "8yo",
          "200yo", "", "",
          "14-32yo", "8yo", "18yo", 
          "10yo"),
  DisturbationType = c("Plantation", "", "Grazing/Logging",
                       "Secondary forest from fallow", "Logging-Thinning-Fire", "",
                       "Plantation",
                       "", "", "",
                       "Logging", "Plantation", "Plantation", 
                       "Plantation")
)
literature %>% 
  select(Effect, Disturbed, Age, DisturbationType, Study, 
         StudyType, Type, Biome, SubBiome) %>% 
  arrange(Effect,
          factor(Disturbed, levels = c("Y", "50%Y", "N", "N?", "?")),
          factor(Age, levels = c("8yo", "10yo", "12yo", "18yo", "1-29yo",
                                 "21yo", "14-32yo", "77yo", "200yo", "600yo", " "))) %>% 
  kable() %>% 
  kable_styling(full_width = F)
```

@@

# References

>>>>

---
title: Biodiversity improves tropical forest resilience
author: Sylvain Schmitt$^{*,1,2}$, Isabelle Maréchaux$^3$, Jerome Chave $^4$, Fabian Fischer$^4$, Camille  Piponiot$^5$, Stéphane Traissac$^6$, Bruno Hérault$^{7,8}$
output:
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: no
    toc: no
    keep_tex: yes
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(kableExtra)
options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
theme_set(bayesplot::theme_default())
opts_chunk$set( echo = F, message = F, warning = F, 
                fig.height = 8, fig.width = 12,
                cache = T, cache.lazy = F)
```

@@

$^1$ CNRS, UMR EcoFoG (Agroparistech, Inra, Université des Antilles, Université de la Guyane, Cirad), Campus Agronomique, 97310 Kourou, French Guiana

$^2$ UMR BIOGECO, Université de Bordeaux, 69 route d’Arcachon, F-33612 Cestas cedex France

$^3$ AMAP, INRA, University of Montpellier, IRD, CIRAD, CNRS, , Montpellier, France

$^4$ Laboratoire Evolution et Diversité Biologique, UMR5174, CNRS, Université Paul Sabatier, IRD, Toulouse Cedex 9, France

$^5$ Université de Guyane, UMR EcoFoG (Agroparistech, CNRS, Inra, Université des
Antilles, Cirad), Kourou, French Guiana

$^6$ AgroParistech, UMR EcoFoG (CNRS, Inra, Université des Antilles, Université de la Guyane, Cirad), Campus Agronomique, 97310 Kourou, French Guiana

$^7$ Cirad, UMR EcoFoG (AgroParistech, CNRS, Inra, Université des Antilles, Université de la Guyane), Campus Agronomique, 97310 Kourou, French Guiana

$^8$ INPHB, Institut National Polytechnique Félix Houphouët-Boigny, BP 1093, Yamoussoukro, Ivory Coast

__$^*$ Corresponding author:__ sylvain.schmitt@agroparistech.fr

\newpage

@@

\newpage

@@

@@

The recovery of the simulated forest at a time $t$, $R(t)$ regarding a given property $X$ was defined as follows (Fig. \@ref(fig:IrGraphFinal)):

@@

When $R(t)$ approaches $R_{eq}=1$, the disturbed and control systems are considered equivalent, indicating a perfect recovery. We then calculated the Euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$ (Fig. \@ref(fig:IrGraphFinal)). Because we set up the intensity of disturbance, our analysis did not include any ecosystem resistance, i.e. the ecosystem capacity to absorb a disturbance event. Consequently, in this particular case, we used the inverse of the cumulative Euclidean distance to the equilibrium as an index of ecosystem resilience, further defined as $I_R$ (Fig. \@ref(fig:IrGraphFinal)):

@@

@@

As species richness ($SR$) increased from 5 to 125 species, functional diversity varied from $0.76\pm0.18$ to $0.96\pm0.02$ and evenness ($FEve$) varied from $0.46\pm0.24$ to $0.64\pm0.02$. Species richness ($SR$) was significantly correlated to functional indices (one-way ANOVA $p-value < 0.001$ for both $FDiv$ and $FEve$) and model $AIC$ was higher without $SR$, consequently we kept only $FDiv$ and $FEve$ within linear regressions. Both functional diversity ($FDiv$) and functional evenness ($FEve$) were positively related with $I_R$, i.e. forest resilience increased with both functional diversity and functional evenness (Fig. \@ref(fig:gIr) and Table \@ref(tab:tgIr)). Similar results were obtained for all disturbance levels, with smaller p-values for higher disturbance intensity.

@@

The net effect of biodiversity on each ecosystem property was quantified and partitioned into selection and complementarity effects after normalizing, as explained in the Methods section. We found that the complementarity effect sharply increased immediately after disturbance, was the dominant effect of biodiversity during the first decades and then progressively vanished (Fig. \@ref(fig:gBE)). When increasing the disturbance intensity, the effect of complementarity increased further and dominated even longer. The selection effect, on the other hand, sharply decreased just after disturbance, and only gradually came to play a more important role, dominating the net biodiversity effect ca. a century after disturbance. Finally, at high disturbance intensity, the net effect of biodiversity on aboveground biomass was still lower than the control 600 years after disturbance (Fig. \@ref(fig:gBE)). We found similar results, but with an amplified signal, for basal area ($BA$) and stem abundance ($N$). Finally, forest growth primary productivity ($GPP$) recovered only within a few years with a dominant complementarity effect.

@@

@@

# References {-}

<div id="refs"></div>

\newpage

## Table captions

__Table \@ref(tab:traits)__ Species-specific parameters used in TROLL from @Marechaux2017 Data originates from the BRIDGE [@Baraloto2010] and TRY [@Kattge2011] datasets.

__Table \@ref(tab:tgIr)__ Effects of functional diversity on forest resilience. Results of the linear regression testing the effect of functional diversity indices (FDiv and FEve) on the resilience index ($I_R$) for forest structure and functioning and for each level of disturbance intensity ($disturb_{intensity}$). To meet the assumption of normality $I_R$ was log-transformed. Test of FDiv and FEve effects was done by linear regression with FDiv and FEve as independent variables. Values of FDiv and FEve are coefficients from linear regressions. Significant $p-values$ are shown in bold. ns, non-significant; . p < 0.1; * p < 0.05; ** p < 0.01; *** p < 0.001.

__Table \@ref(tab:tNE)__ Mean and standard deviation of the biodiversity net effect across mature simulations before the disturbance event (i.e. after 600 years of regeneration) for different ecosystem variables.

## Figure captions

__Figure \@ref(fig:IrGraphFinal)__ Definition of recovery function R and resilience index $I_R$ shown for aboveground biomass dynamics for a simulation with $disturb_{intensity}=50\%$.

__Figure \@ref(fig:gIr)__ Variation of forest resilience with taxonomic and functional diversity for different levels of disturbance. Resilience index $I_R$ was represented against functional diversity ($FDiv$) for different levels of disturbance ($disturb_{intensity}$} = 25, 50 and 75% of total basal area); dot shapes varies with species richness whereas dot color indicates functional evenness ($FEve$). Solid lines and grey areas show fitted relationships with their mean and variance of the form $I_R = FDiv + \epsilon$.

__Figure \@ref(fig:gBE)__ Post-disturbance trajectories of complementarity and selection effects on aboveground biomass for increasing levels of disturbance (25%, 50% and 75% of basal area loss). The net biodiversity ($NE$), complementarity ($CE$) and selection ($SE$) effects were normalized by the net effect of biodiversity in the control simulation ($disturb_{intensity}=0\%$). Solid lines represents the mean value whereas shaded areas represent the range of values across the 20 simulations with a species richness ($SR$) of 125. Time axis has been square-root transformed to highlight first decades dynamic.

\newpage

<!-- tables -->

```{r tgIr}
Ir %>% 
  left_join(doe, by = "sim") %>% 
  ungroup() %>% 
  mutate(SR = as.numeric(SR)) %>% 
  select(disturbance, structure, production,
         SR, FEve, FDiv) %>% 
  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
                 variable.name = "Ir") %>% 
  group_by(Ir, disturbance) %>%
  do(fit = lm(log(value) ~ FEve + FDiv, data = .)) %>% 
  tidy(fit) %>% 
  ungroup() %>% 
  filter(term != "(Intercept)") %>% 
  filter(!grepl("SR", term)) %>% 
  mutate(`P-value` = ifelse(p.value < 0.001, "***",
                            ifelse(p.value < 0.01, "**",
                                   ifelse(p.value < 0.05, "*",
                                          ifelse(p.value < 0.1, ".", 
                                                 "n.s."))))) %>% 
  mutate(Coefficient = round(estimate,2)) %>% 
  mutate(value = paste(Coefficient, `P-value`)) %>%
  mutate(value = cell_spec(value,
                           bold = ifelse(grepl("*", value,
                                               fixed = T), T, F))) %>%
  reshape2::dcast(Ir + disturbance ~ term, value.var = "value") %>% 
  left_join(Ir %>% 
              left_join(doe, by = "sim") %>% 
              ungroup() %>% 
              mutate(SR = as.factor(SR)) %>% 
              select(disturbance, structure, production,
                     SR, FEve, FDiv) %>% 
              reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
                             variable.name = "Ir") %>% 
              group_by(Ir, disturbance) %>%
              do(fit = lm(log(value) ~ FEve + FDiv, data = .)) %>% 
              glance(fit) %>% 
              mutate(r.squared = round(r.squared, 2)) %>% 
              select(Ir, disturbance, r.squared)) %>% 
  mutate(Ir = ifelse(Ir == "production", "functioning", "structure")) %>% 
  kable(caption = "Effects of functional diversity on forest resilience. Results of the linear regression testing the effect of functional diversity indices (FDiv and FEve) on the resilience index ($I_R$) for forest structure and functioning and for each level of disturbance intensity ($disturb_{intensity}$). To meet the assumption of normality $I_R$ was log-transformed. Test of FDiv and FEve effects was done by linear regression with FDiv and FEve as independent variables. Values of FDiv and FEve are coefficients from linear regressions. Significant $p-values$ are shown in bold. ns, non-significant; . p < 0.1; * p < 0.05; ** p < 0.01; *** p < 0.001.",
        col.names = c("$I_R$", "Disturbance intensity",
                      "FDiv", "FEve", "$R^2$"), escape = F, format = "pandoc") %>% 
  kable_styling(full_width = F) %>% 
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1:2)
```

<!-- figures -->

```{r gIr, echo=FALSE, fig.cap='Variation of forest resilience with taxonomic and functional diversity for different levels of disturbance. Resilience index $I_R$ was represented against functional diversity ($FDiv$) for different levels of disturbance ($disturb_{intensity}$} = 25, 50 and 75% of total basal area); dot shapes varies with species richness whereas dot color indicates functional evenness ($FEve$). Solid lines and grey areas show fitted relationships with their mean and variance of the form $I_R = FDiv + \\epsilon$.', fig.height=12, dpi=600}
Ir %>% 
  left_join(doe, by = "sim") %>% 
  ungroup() %>% 
  select(disturbance, 
         structure, SR, FEve, FDiv) %>% 
  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), value.name = "Ir") %>% 
  ggplot(aes(FDiv, Ir, shape = SR, col = FEve, group = NA)) +
  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
  geom_point(size = 4) +
  facet_wrap(~disturbance, scales = "free", labeller = 'label_both', 
             nrow = 3) + 
  xlab('Functional diversity (FDiv)') +
  ylab('Resilience indice Ir') +
  scale_shape_discrete('Species\nrichness\n(SR)') +
  scale_color_continuous('Functional\nevenness\n(FEve)') +
  theme(text = element_text(size=20),
        legend.key.size=unit(4,"line"))
```

```{r gBE, echo=FALSE, fig.cap='Post-disturbance trajectories of complementarity and selection effects on aboveground biomass for increasing levels of disturbance (25%, 50% and 75% of basal area loss). The net biodiversity ($NE$), complementarity ($CE$) and selection ($SE$) effects were normalized by the net effect of biodiversity in the control simulation ($disturb_{intensity}=0\\%$). Solid lines represents the mean value whereas shaded areas represent the range of values across the 20 simulations with a species richness ($SR$) of 125. Time axis has been square-root transformed to highlight first decades dynamic.', fig.height=12, dpi=600}
BEresilience %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  # filter(BE %in% c('SE', 'CE')) %>%
  group_by(time, BE, disturbance, SR) %>%
  summarise(qt1 =quantile(agb, probs=0.05),
            qt2 =quantile(agb, probs=0.75),
            mean=mean(agb),
            n=n()) %>% 
  ggplot(aes(x = time)) +
  geom_hline(yintercept = 1, color = 'red') +
  geom_hline(yintercept = 0, color = 'blue') +
  geom_ribbon(aes(ymin = qt1, ymax = qt2, fill = BE), alpha = 0.3) +
  geom_line(aes(y = mean, color = BE)) +
  facet_grid(disturbance ~ SR, labeller = 'label_both') +
  xlab('Time since disturbance (years)') +
  ylab('Biodiversity effect on aboveground biomass') +
  scale_x_sqrt() +
  theme(text = element_text(size=20),
        legend.key.size=unit(2,"line"))
```
